SDSS-IV MaNGA: The Radial Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2021-01-21 maintex.tex
revised iniside-IFU sample section

{\bf 
We use {\sc spfit} to model the MaNGA datacubes. The IDL package is publicly available\footnote{\url{https://github.com/fuhaiastro/spfit}} and was first used in our previous study of close galaxy pairs in MaNGA \citepalias{Fu:2018}. While data products from the MaNGA Data Analysis Pipeline \citep[DAP;][]{Belfiore:2019} and {\sc PIPE3D} \citep{Sanchez:2016a, Sanchez:2016b} are available, {\sc spfit} allows us to combine spectra inside a given aperture before fitting the combined spectrum. The feature alone made it better suited for this project. Additional features of {\sc spfit} include: (1) simultaneously fitting emission lines and stellar continuum, (2) coadding spaxels with either a Voronoi tessellation or arbitrary polygons while accounting for covariances, (3) modeling asymmetric emission lines, broad AGN emission lines, AGN continuum, and various dust-extinction laws, and (4) utilizing multi-threading when processing a large number of datacubes.

@@

\begin{table}
\begin{center}
\caption{Close Galaxy Pairs and Multiples in MaNGA IFUs}
\label{tab:coords}
\begin{tabular}{lc ccc}
\hline
\hline
Plate-IFU & Index & R.A. (J2000) & Decl. (J2000) & Redshift \\
          &           & (deg)        & (deg)         &     \\
\hline
 7443-12703&0&229.5255758&+42.7458538&0.04043\\
    \nodata&1&229.5265348&+42.7440666&0.04079\\
  7958-3701&0&257.5338400&+33.5989046&0.11015\\
    \nodata&1&257.5344200&+33.5984600&0.10920\\
    \nodata&2&257.5326566&+33.5979470&0.11031\\
  7960-3701&0&257.0857469&+31.7469109&0.10819\\
    \nodata&1&257.0850100&+31.7470300&0.10800\\
  7960-3702&0&258.2893316&+31.5820375&0.02961\\
    \nodata&1&258.2913540&+31.5812244&0.02943\\
  7962-3702&0&259.0637432&+28.0140065&0.10725\\
    \nodata&1&259.0639447&+28.0130439&0.10753\\
\hline
\end{tabular}
\end{center}
\tablecomments{
This table lists a total of 404 plate-IFUs that contain close galaxy pairs and multiples in the eighth MaNGA Product Launch (MPL-8). The second column lists the component index, where ``0'' indicates the primary target of the MaNGA observation. For each IFU, all components within $\pm$2000\,km~s$^{-1}$ of the primary are listed, sorted in ascending angular distance from the primary.\\
(This table is available in its entirety in a machine-readable form in the online journal. A portion is shown here for guidance regarding its form and content.)
}
\end{table}

@@

To select paired galaxies out of our object catalog, we inspect the spectra of each object. The spectra of the identified objects is extracted through a 1\arcsec\ circular aperture and fitted assuming the MaNGA target's redshift and then manually sorted into the following categories: ``good" galaxy spectra, broad-line AGN, foreground star, foreground/background galaxies, or poor S/N objects. The ``good" galaxy spectra are the objects whose spectra are well modeled by {\sc spfit} at the target galaxy's redshift, whether it is the target galaxy itself or a nearby companion galaxy. This means that the companion galaxy can be within approximately $\pm$3000 km s$^{-1}$ of the MaNGA target. We found 6573 ``good" objects, 57 broad-line AGN, 836 foreground stars, 319 foreground/background galaxies, and 1546 objects with poor S/N.

@@

Among the 6142 MaNGA IFUs considered here, 404 have multiple galaxy objects within the IFU. We further restrict the sample by setting a relative velocity cut of $\Delta v$ $<$ 500 km s$^{-1}$ to remove projected companions. Given the redshift range of the MaNGA sample and the size of MaNGA's IFUs, the maximum projected separation for a companion galaxy in the IFU is $\sim$40 kpc. Again, the galaxy sample is also restricted to galaxies with a stellar mass range of \logm\ $=$ 9.0$-$11.5 and galaxies in the Primary or Secondary MaNGA subsamples. We also require that the galaxies are classified as star forming as described in \S~\ref{sec:data}. These requirements reduce our sample down to 54 star forming MaNGA targets with inside-IFU companions. \textbf{While we use the identified companions to select galaxy pairs, we only build radial profiles for the MaNGA target galaxies. We do this because the identified companions are not included in the NSA catalog which we use to build the radial profiles for the galaxies.}

Several of the target galaxies in our sample have multiple companions. From the whole set of 404 IFUs, there are 327 pairs, 67 triplets, 7 quadruplets, and 1 quintuplet. In cases where a MaNGA target has multiple viable companions galaxies, we chose the closest companion when measuring the pair's mass ratio and projected separation.  \textbf{We show the whole 404 inside-IFU pair sample in Table \ref{tab:coords}.}

@@

\subsection{Specific Star Formation Rate}
\textbf{We calculate the star formation rates in the pair and control galaxies with the emission lines extracted with our spectra fitting code, {\sc SPFIT} (described in Section \ref{sec:spfit}).} We correct the emission lines for reddening using the extinction curve from \citet{Cardelli:1989} with updated coefficients from \citet{ODonnell:1994}. The extinction is parameterized as $R_V$ $\equiv$ $A_V/E(B-V)$ $=$ 3.1, where we estimate the value of the $V$-band extinction, $A_V$, by comparing the H$\alpha$/H$\beta$ ratio to the expected value of 2.85 for Case-B recombination.

@@

\textbf{
We check our measurement of the specific star formation rate with the equivalent width of the H$\alpha$ line, \ewha, since it is a known proxy for the sSFR. This is useful as the sSFR is dependent on {\sc SPFIT}'s measurement of the stellar mass while \ewha\ is an observable. As we will show in Section \ref{sec:results}, the radial profiles of \ewha\ is consistent with radial profiles of the sSFR.
}

@@

\textbf{ 
We repeat this analysis for the \ewha\ in Figure \ref{fig:ewha_prof}. We find that the \ewha\ profiles are in close agreement with the sSFR profiles, which shows that we can be confident of {\sc SPFIT}'s measurement of the sSFR. We do see that the pair-control offsets are lightly suppressed using the \ewha\ as the $\Delta$log(\ewha) is only enhanced by 0.25 dex where the $\Delta$log(sSFR) was enhanced by 0.30 dex. 
}

>>>>

{\bf 
We use {\sc spfit} to model the MaNGA datacubes. The IDL package is publicly available\footnote{\url{https://github.com/fuhaiastro/spfit}} and was first used in our previous study of close galaxy pairs in MaNGA \citepalias{Fu:2018}. While data products from the MaNGA Data Analysis Pipeline \citep[DAP;][]{Belfiore:2019} and {\sc pipe3d} \citep{Sanchez:2016a, Sanchez:2016b} are available, {\sc spfit} allows us to combine spectra inside a given aperture before fitting the combined spectrum. The feature alone made it better suited for this project. Additional features of {\sc spfit} include: (1) simultaneously fitting emission lines and stellar continuum, (2) coadding spaxels with either a Voronoi tessellation or arbitrary polygons while accounting for covariances, (3) modeling asymmetric emission lines, broad AGN emission lines, AGN continuum, and various dust-extinction laws, and (4) utilizing multi-threading when processing a large number of datacubes.

@@

\begin{table}
\begin{center}
\caption{Close Galaxy Pairs and Multiples in MaNGA IFUs}
\label{tab:sample}
\begin{tabular}{lc ccc}
\hline
\hline
Plate-IFU & Index & R.A. (J2000) & Decl. (J2000) & Redshift \\
          &           & (deg)        & (deg)         &     \\
\hline
 7443-12703&0&229.5255758&+42.7458538&0.04043\\
    \nodata&1&229.5265348&+42.7440666&0.04079\\
  7958-3701&0&257.5338400&+33.5989046&0.11015\\
    \nodata&1&257.5344200&+33.5984600&0.10920\\
    \nodata&2&257.5326566&+33.5979470&0.11031\\
  7960-3701&0&257.0857469&+31.7469109&0.10819\\
    \nodata&1&257.0850100&+31.7470300&0.10800\\
  7960-3702&0&258.2893316&+31.5820375&0.02961\\
    \nodata&1&258.2913540&+31.5812244&0.02943\\
  7962-3702&0&259.0637432&+28.0140065&0.10725\\
    \nodata&1&259.0639447&+28.0130439&0.10753\\
\hline
\end{tabular}
\end{center}
\tablecomments{
This table lists a total of 404 plate-IFUs that contain close galaxy pairs and multiples in the eighth MaNGA Product Launch (MPL-8). The second column lists the component index, where ``0'' indicates the primary target of the MaNGA observation. For each IFU, all components within $\pm$2000\,km~s$^{-1}$ of the primary are listed, sorted in ascending angular distance from the primary.\\
(This table is available in its entirety in a machine-readable form in the online journal. A portion is shown here for guidance regarding its form and content.)
}
\end{table}

@@

To select paired galaxies out of our object catalog, we inspect the spectra of each object. The spectra of the identified objects is extracted through a 1\arcsec\ circular aperture and fitted assuming the MaNGA target's redshift and then manually sorted into the following categories: ``good" galaxy spectra, broad-line AGN, foreground star, foreground/background galaxies, or poor S/N objects. The ``good" galaxy spectra are the objects whose spectra are well modeled by {\sc spfit} at the target galaxy's redshift, whether it is the target galaxy itself or a nearby companion galaxy. This means that the companion galaxy can be within approximately $\pm$2000 km s$^{-1}$ of the MaNGA target. We found 6573 ``good" objects, 57 broad-line AGN, 836 foreground stars, 319 foreground/background galaxies, and 1546 objects with poor S/N.

@@

{\bf Among the 6142 MaNGA IFUs considered here, 404 cover close galaxy pairs and multiples. The sample includes 327 pairs, 67 triplets, 7 quadruplets, and 1 quintuplet. Table\,\ref{tab:sample} lists the coordinates and redshifts of these galaxy components in pairs/multiples.} When a MaNGA target has multiple companions galaxies, we chose the closest companion to define the pair's mass ratio and projected separation.

For this study, we further restrict the sample by setting a relative velocity cut of $\Delta v$ $<$ 500 km s$^{-1}$. Given the redshift range of the MaNGA sample and the size of MaNGA's IFUs, the maximum projected separation for a companion galaxy in the IFU is $\sim$40 kpc. Again, the galaxy sample is also restricted to galaxies with a stellar mass range of \logm\ $=$ 9.0$-$11.5 and galaxies in the Primary or Secondary MaNGA subsamples. We also require that the galaxies are classified as star forming as described in \S~\ref{sec:data}. These requirements reduce our sample down to 54 star forming MaNGA targets with inside-IFU companions. \textbf{While we use the identified companions to select galaxy pairs, we only build radial profiles for the MaNGA target galaxies. We do this because the identified companions are not included in the NSA catalog which we use to build the radial profiles for the galaxies.}

@@

\subsection{Specific Star Formation Rate}
\textbf{We calculate the star formation rates in the pair and control galaxies with the emission lines extracted with our spectra fitting code, {\sc spfit} (described in Section \ref{sec:spfit}).} We correct the emission lines for reddening using the extinction curve from \citet{Cardelli:1989} with updated coefficients from \citet{ODonnell:1994}. The extinction is parameterized as $R_V$ $\equiv$ $A_V/E(B-V)$ $=$ 3.1, where we estimate the value of the $V$-band extinction, $A_V$, by comparing the H$\alpha$/H$\beta$ ratio to the expected value of 2.85 for Case-B recombination.

@@

\textbf{
We check our measurement of the specific star formation rate with the equivalent width of the H$\alpha$ line, \ewha, since it is a known proxy for the sSFR. This is useful as the sSFR is dependent on {\sc spfit}'s measurement of the stellar mass while \ewha\ is an observable. As we will show in Section \ref{sec:results}, the radial profiles of \ewha\ is consistent with radial profiles of the sSFR.
}

@@

\textbf{ 
We repeat this analysis for the \ewha\ in Figure \ref{fig:ewha_prof}. We find that the \ewha\ profiles are in close agreement with the sSFR profiles, which shows that we can be confident of {\sc spfit}'s measurement of the sSFR. We do see that the pair-control offsets are lightly suppressed using the \ewha\ as the $\Delta$log(\ewha) is only enhanced by 0.25 dex where the $\Delta$log(sSFR) was enhanced by 0.30 dex. 
}
